---
title: "Shiny RMarkdown Assigment"
---

# Shiny Apps and RMarkdown

In this assignment, I am going to explore the **nyc_dop_juvenile_rearrest_rate dataset** to test my coding skills with shiny apps and RMarkdown.

## Loading my libraries

```{r}
#| label: loading-libraries
#| message: false
#| warning: true
#| paged-print : false

library(shiny)
library(tidyverse)
library(nycOpenData)
library(DT)
```

## Calling my data

Let's bring the data for the assignment.

```{r}
#| label: loading-dataset
#| message: false
#| warning: true
#| paged-print : false

data_rr <- nyc_dop_juvenile_rearrest_rate(limit = 10000)
head(data_rr)
```

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("month", "Choose a Month:",
         choices = c("January", "February",   "March", "April", "May", "June", "July", 
  "August", "September", "October", 
  "November", "December")),
  sliderInput("top_n", "Year:", 2015, 2020, 2025)
)

renderPlot({
  data_rr %>%
        filter(month == input$month) %>%
        count(rate, sort = T) %>%
        slice_head(n=input$top_n) %>%
        ggplot(aes(n,fct_reorder(rate, n))) +
        geom_col(fill = "lightblue")
    })
```
